LaMachine: A meta-distribution for NLP software

### 2019-01-29 docs/lamachine_clarin2018.tex
fixed the broken sentence, added note to case study 2 and extended the acknowlegdements.

    \item The software must have a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository. \footnote{e.g. Github, Gitlab, Bitbucket, provided
        the respository is public}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item Participating software must be actively maintained (i.e. not outdated or abandoned) and not place any demands on outdated
        dependencies. %verwerkt <-- maarten help even met formuleren hier, ik bedoel dat zoiets als Pattern dus helaas niet mee kan doen omdat het niet meer onderhouden wordt }only maintained/uptodate date software that can compile using the current version of compilers and dependencies/libraries
\end{enumerate}

@@

LaMachine exists since May 2015 and has been used extensively ever since by numerous users. In early 2018 version 2 was
released which was a significant redesign, powered by Ansible.

@@

The ICT-Research Platform Netherlands and NWO organise a yearly one-week workshop `ICT with Industry'
\footnote{https://ict-research.nl/ict-with-industry/ictwi2019/} to stimulate collaboration between industry and
academia. The industrial partner provides a problem and a team of researchers from different backgrounds and
universities collaborate to come up with solutions. We participated in the 2019 edition on the case study by the Dutch
Royal Library who %@Iris: zin niet afgemaakt?

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining.  However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: For team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine and we did not manage to get X-forwarding working in the
Ubuntu Linux VM. After a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of LaMachine on certain aspects is beyond the reach of a data scientist
without system administration skills. But this certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills.
 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?

@@

This research was funded by NWO CLARIN-NL, CLARIAH and the ZonMw project {\it Tekstmining in het toezicht: een exploratieve studie naar de automatische verwerking van klachten ingediend bij het Landelijk Meldpunt Zorg}, project number 516004614. We thank all project partners: the Dutch Health Inspectorate, IQhealthcare, and Tim Voets for their valuable contributions and help in the ZonMw project.

>>>>

    \item The software must have a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository.\footnote{e.g. Github, Gitlab, Bitbucket, provided
        the respository is public}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item Participating software must be actively maintained (i.e. not outdated or abandoned) and not place any demands on outdated
        dependencies. %verwerkt <-- maarten help even met formuleren hier, ik bedoel dat zoiets als Pattern dus helaas niet mee kan doen omdat het niet meer onderhouden wordt }only maintained/uptodate date software that can compile using the current version of compilers and dependencies/libraries
\end{enumerate}

@@

LaMachine exists since May 2015 and has been used extensively ever since by numerous users. In early 2018 version 2 was
released which was a significant redesign, powered by Ansible\footnote{\url{https://www.ansible.com}}.

@@

The ICT-Research Platform Netherlands and NWO organise a yearly one-week workshop `ICT with Industry'
\footnote{https://ict-research.nl/ict-with-industry/ictwi2019/} to stimulate collaboration between industry and
academia. The industrial partner provides a problem and a team of researchers from different backgrounds and
universities collaborate to come up with solutions. We participated in the 2019 edition on the case study by the Dutch Royal Library who wanted to investigate automatic methods for cataloguing of textual cultural heritage objects, in this particular case a large collection of digital dissertations.

@@

LaMachine offered a convenient platform for a range of different explorations and experiments in the area of NLP and
text mining.  However, for some situations LaMachine, or rather Linux in general, was not a good fit for the audience of
the workshop: For team members who did not have experience with a non-Windows environment, LaMachine was not a suitable
or useful tool. The limit of LaMachine was also reached for members who wanted to use desktop text editors with a
graphical user interface as this is not offered by LaMachine and we did not manage to get X-forwarding working in the
Ubuntu Linux VM. After a few attempts the team gave up on resolving this issue due to time pressure. This, also
demonstrates that fine-tuning the configuration of LaMachine on certain aspects is beyond the reach of a data scientist
without system administration skills. But this certainly also applies also to the installation as a whole in the
SURFsara context, which involved things like the partitioning, formatting and mounting of (virtual) drives and setting
up user accounts on the shared VM, all of which require some system administration skills.
Lamachine was convenient and speeded up writing code as the most common scientific data-related packages are already present in Lamachine.

 %verwerkt <-- <iris> it moet nog wat beter geformuleerd worden maar je snaprt wat ik bedoel
%@iris: hebben jullie verder geen gebruik proberen te maken van de Jupyter Lab omgeving? Dat biedt in ieder geval iets
%grafisch wat via web te benaderen is?
%@maarten helaas lukte dat ook niet, dat liep bij mij steeds vast. - ik heb opmerking van alex over tijdswinst toegevoefd

@@

This research was funded by NWO CLARIN-NL, CLARIAH and the ZonMw project {\it Tekstmining in het toezicht: een exploratieve studie naar de automatische verwerking van klachten ingediend bij het Landelijk Meldpunt Zorg}, project number 516004614. We thank the Dutch Health Inspectorate, IQhealthcare, and Tim Voets for their valuable contributions and help in the ZonMw project, and NWO, the ICT Research Platform Nederland, the Lorentz Centre, the Royal Library and all our team members for making the ICT with Industry workshop a success.
